---
title: "SimDraft"
author: '""'
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(moments)
library(SimMultiCorrData)
options(scipen = 99999)
library(PearsonDS)
library(lme4)
library(lmerTest)

child_data <- readRDS(here::here("child_data.Rds"))
```

###### LME Model w/ adolescent data

.AID = identifier
.decisions_eat = do parent's give child freedom to decide what they eat (0 = no, 1 = yes)
.run_away = ever run away? (0 = no, 3 = yes)
.cig = ever smoked cigarettes? (0 = no, 1 = yes)
.physical_fight = been in a fight? (0 = no, 1 = yes)
.alcohol = ever had alcohol? (0 = no, 1 = yes)
.wave = wave of study 
```{r}
glimpse(child_data)
```

Intercept only model
```{r}
intercept_model <- lmer(bmi ~  decisions_eat +  run_away + cig + alcohol + physical_fight + real_age + (1|AID), asian_women_ds) 
summary(intercept_model)
```

Intercept and slope model

```{r}
intercept_slope_model <- lmer(bmi ~  decisions_eat +  run_away + cig + alcohol + physical_fight + real_age + wave + ( wave |AID), asian_women_ds) 
summary(intercept_slope_model)
print(VarCorr(intercept_slope_model),comp=c("Variance","Std.Dev."),digits=2)
```

# SimMultiCorrData

```{r}

fitdistr(child_data$bmi,densfun = "log-normal") # Assuming log-normal, find parameters
calc_moments(asian_women_ds$bmi) # gives skewness + kurtosis vals based on data 
m<-round(calc_theory(Dist = "Lognormal", params = c(0,.5)),8) # gives kurt + skew based on parameters from line 53
corr.data <- c(1,-.38,-.38,1)
corr1 <- matrix(corr.data,nrow=2)

valid_corr(n = 12, k_cont = 2, method = "Fleishman", means = c(m[1],m[1]), vars = c(m[2],m[2]), skews = c(m[3],m[3]),
           skurts =  c(m[4],m[4]), rho = corr1)

r_effects <-rcorrvar(n = 12, k_cont = 2, method = 'Fleishman', means = c(m[1],m[1]), vars = c(m[2],m[2]), skews = c(m[3],m[3]),
           skurts =  c(m[4],m[4]),
          errorloop = TRUE, rho = corr1
         )

```




###### SIMULATION

```{r}
set.seed(15)
n.indiv <- 12
n.measurements <-4
```

# Random Intercept Model, using skewness + kurtosis values from Arnau paper lognormal distribution 

Generating values for intercept random effect + error term
```{r}
nonnormal_vals <- nonnormvar1("Fleishman", means = 0, vars = 1, skews =  1.75,
                    skurts = 5.9, , n = n.indiv, seed = 1234)$continuous_variable$V1

hist(nonnormal_vals)

errors <- rnorm(nrow(dt), 0, 2)
```

```{r}
# setting up sim dataset w/ n.indiv subjects that each have n.measurements data points at different time points 

data <- expand.grid(indiv = LETTERS[1:n.indiv], time = 1:n.measurements)
matrix_fixedeffects <- model.matrix(~ time, data)   # model matrix for fixed effects

betas <- c(3.1, 0)   # fixed effects coefficient vector, true val of slope is 0 

matrix_randeffects <- model.matrix(~ 0 + indiv, data)   # model matrix for random intercepts


# generating Y values
data$Y_val <- matrix_fixedeffects %*% betas + matrix_randeffects %*% nonnormal_vals + errors


# construct model
intercept_m0 <- lmer(Y_val ~ 1 +indiv + time + (1|indiv) , data) 
summary(intercept_m0)
anova(intercept_m0)
VarCorr(intercept_m0)
```


# Random intercept + slope model?

```{r}

data <- expand.grid(indiv = LETTERS[1:n.indiv], time = 1:n.measurements)
matrix_fixedeffects <- model.matrix(~ time, data)   # model matrix for fixed effects

betas <- c(3.1, 0)   # fixed effects coefficient vector

matrix_intercept <- model.matrix(~ 0 + indiv, data)   # model matrix for random intercepts

matrix_slope <-  model.matrix(~ 0 + indiv, data) * data$time   # model matrix for random slopes

r_effects1 <- r_effects$continuous_variables$V1
r_effects2 <- r_effects$continuous_variables$V2
errors <- rnorm(nrow(data), 0, 2)  

data$Y_val <- matrix_fixedeffects %*% betas + matrix_intercept %*% r_effects1 + matrix_slope %*% r_effects2 + errors

# construct model
intercept_m1 <- lmer(Y_val ~ time + (time|indiv) , data) 
summary(intercept_m1)
anova(intercept_m1)
VarCorr(intercept_m1)

```

