---
title: "SimDraft"
author: '""'
date: "1/28/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(MASS)
library(moments)
library(SimMultiCorrData)
library(arm)
options(scipen = 99999)
library(PearsonDS)
library(lme4)
library(lmerTest)
library(pbkrtest)

child_data <- readRDS(here::here("child_data.Rds"))
```

###### LME Model w/ adolescent data

.AID = identifier
.decisions_eat = do parent's give child freedom to decide what they eat (0 = no, 1 = yes)
.run_away = ever run away? (0 = no, 3 = yes)
.cig = ever smoked cigarettes? (0 = no, 1 = yes)
.physical_fight = been in a fight? (0 = no, 1 = yes)
.alcohol = ever had alcohol? (0 = no, 1 = yes)
.wave = wave of study 
```{r}
glimpse(child_data)
```

Intercept only model
```{r}
reg_model <- lm(bmi ~  decisions_eat +  run_away + cig + alcohol + physical_fight + real_age  + wave, child_data)
intercept_model <- lmer(bmi ~  1+ decisions_eat +  run_away + cig + alcohol + physical_fight + real_age + (1|AID) + wave, child_data) 
summary(intercept_model,ddf = "Kenward-Roger")
summary(intercept_model,ddf = "Satterthwaite")
summary(reg_model)
hist(child_data$bmi)
tidy(anova(intercept_model), ddf = "Kenward-Roger")
tidy(anova(intercept_model), ddf = "Satterthwaite")
vc <- data.frame(VarCorr(intercept_model))
var_rand_intercept<- vc[1,4]
var_rand_residual <- vc[2,4]
ranef(intercept_model)

plot(intercept_model)
qqnorm(intercept_model, ~ranef(., level=0))
qqnorm(resid(intercept_model))
abline(0,1, col = "red", lty = 2)
qqnorm(resid(reg_model))
abline(0,1, col = "red", lty = 2)
plot(ranef(intercept_model, level = 1))

qqmath(ranef(intercept_model, postVar = TRUE), strip = FALSE)
```

Intercept and slope model

```{r}
child_data <- child_data %>%
  mutate(wave = as.factor(wave))
intercept_slope_model <- lmer(bmi ~  decisions_eat +  run_away + cig + alcohol + physical_fight + real_age + wave + ( 1 + wave|AID), child_data) 
summary(intercept_slope_model, ddf ="Kenward-Roger" )
summary(intercept_slope_model, ddf = "Satterthwaite")
print(VarCorr(intercept_slope_model),comp=c("Variance","Std.Dev."),digits=2)
glimpse(child_data)
str(summary(intercept_slope_model))
ranova(intercept_model)
se.ranef(intercept_slope_model)
vc <- data.frame(VarCorr(intercept_slope_model))
var_rand_intercept<- vc[1,4]
var_rand_slope <- vc[2,4]
var_rand_residual <- vc[4,4]

```

# SimMultiCorrData

```{r}

fitdistr(child_data$bmi,densfun = "log-normal") # Assuming log-normal, find parameters
calc_moments(child_data$bmi) # gives skewness + kurtosis vals based on data 
m<-round(calc_theory(Dist = "Lognormal", params = c(0,.5)),8) # gives kurt + skew based on parameters from line 53
corr.data <- c(1,-.38,-.38,1)
corr1 <- matrix(corr.data,nrow=2)

m[4]

valid_corr(n = 12, k_cont = 2, method = "Fleishman", means = c(m[1],m[1]), vars = c(m[2],m[2]), skews = c(m[3],m[3]),
           skurts =  c(m[4],m[4]), rho = corr1)

r_effects <-rcorrvar(n = 12, k_cont = 2, method = 'Fleishman', means = c(m[1],m[1]), vars = c(m[2],m[2]), skews = c(m[3],m[3]),
           skurts =  c(m[4],m[4]),
          errorloop = TRUE, rho = corr1)


r_effects <-rcorrvar(n = 12, k_cont = 1, method = 'Fleishman', means = m[1], vars = m[2], skews = m[3],
           skurts =  m[4],
          errorloop = TRUE
         )
        
r_effects

```

## generate graphs of the Non normal distributions (lognormal),just to show



###### SIMULATION

```{r}
set.seed(15)
n.indiv <- 26
n.measurements <-4
```

# Random Intercept Model, using skewness + kurtosis values from Arnau paper lognormal distribution 

Generating values for intercept random effect + error term
```{r}
nonnormal_vals <- nonnormvar1("Fleishman", means = m[1], vars = m[2]^2, skews =  m[3],
                    skurts = m[4], , n = n.indiv, seed = 1234)$continuous_variable$V1

hist(nonnormal_vals)

test <- c(1)
test[2]

errors <- rnorm(nrow(dt), 0, 2)
```

```{r}
# setting up sim dataset w/ n.indiv subjects that each have n.measurements data points at different time points 

data <- expand.grid(indiv = LETTERS[1:n.indiv], time = 1:n.measurements)
matrix_fixedeffects <- model.matrix(~ time, data)   # model matrix for fixed effects

betas <- c(3.1, 0)   # fixed effects coefficient vector, true val of slope is 0 

matrix_randeffects <- model.matrix(~ 0 + indiv, data)   # model matrix for random intercepts


# generating Y values
data$Y_val <- matrix_fixedeffects %*% betas + matrix_randeffects %*% nonnormal_vals + errors


# construct model
intercept_m0 <- lmer(Y_val ~ 1 + time + (1|indiv) , data) 
summary(intercept_m0)
str(anova(intercept_m0))
anova(intercept_m0)$`DenDF`
VarCorr(intercept_m0)
```


# Random intercept + slope model?

```{r}

data <- expand.grid(indiv = LETTERS[1:n.indiv], time = 1:n.measurements)
matrix_fixedeffects <- model.matrix(~ time, data)   # model matrix for fixed effects

betas <- c(3.1, 0)   # fixed effects coefficient vector

matrix_intercept <- model.matrix(~ 0 + indiv, data)   # model matrix for random intercepts

matrix_slope <-  model.matrix(~ 0 + indiv, data) * data$time   # model matrix for random slopes

r_effects1 <- r_effects$continuous_variables$V1
r_effects2 <- r_effects$continuous_variables$V2
errors <- rnorm(nrow(data), 0, 2)  

data$Y_val <- matrix_fixedeffects %*% betas + matrix_intercept %*% r_effects1 + matrix_slope %*% r_effects2 + errors

# construct model
intercept_m1 <- lmerTest::lmer(Y_val ~ time + (time|indiv) , data) 
summary(intercept_m1)
anova(intercept_m1)$`Pr(>F)`
VarCorr(intercept_m1)
anova(intercept_m1,ddf = "Satterthwaite")$DenDF

anova(intercept_m1, ddf = "Kenward-Roger")
anova(intercept_m1, ddf = "Satterthwaite")
lmerTest::ranova()
```

