---
title: "Untitled"
author: '""'
date: "8/31/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(mosaic)
```


```{r}
load(file = "~/Desktop/ICPSR_36692/DS0002/36692-0002-Data.rda")
load(file = "~/Desktop/ICPSR_36692/DS0001/36692-0001-Data.rda")
load(file = "~/Desktop/ICPSR_36692/DS0003/36692-0003-Data.rda")

data20 <- da36692.0002

data98 <- da36692.0001

data02 <- da36692.0003
```

```{r}
data20 <- data20 %>%
  select(ID,A52,A53A1,B11,B21,B22,B23,B24 ,B25,B26,F103A4,F92A5,TYPE)

data98 <- data98 %>%
  select(ID,A52,A53A1,B11,B21,B22,B23,B24 ,B25,B26,F101A4,F92A5)

data02 <- data02 %>%
  select(ID,A52,A53A1,B11,B21,B22,B23,B24 ,B25,B26,F103A4,F92A5)


inner_join_test <- data20 %>%
  inner_join(data98, by = c("ID"= "ID")) %>%
  inner_join(data02, by = c("ID"= "ID"))

```


Variables of interest:
R in front of variable means proxy

F103A4 - Frequent visits by kids

PROV - province
TYPE - type of elder (if they were a replacement to deceased elder or new addition)
TRUEAGE - 
A1 - sex
A41 - province
A43 - urban or rural
A52 - how many people living with you
A53A1 - relationship between 1st person living with you
B11 - self reported quality of life
B12- self-reported health
B21- looking at the bright side of things
B22 - keep belongings neat and clean
B23- fearful / anxious
B24 - lonely and isolated
B25 - make own decision
B26 - useless with age
B27 - Are you as happy as when you were younger

```{r}
tally(data20$B11)
```


C11 - C16 - cognitive testing

```{r}
tally(data$C11)


data %>%
  select(C11, C12,C13,C14,C15,C16)

tally(data$B12)
tally(data$B21)

tally(data$D73)
```

A54A - year being admitted to nursing home

FOODS :

D1 - staple food
D31 - Fresh fruits
D32- vegetables
fish, eggs, meat, 


D73 - age started smoking
D83 - age started drinking : 83 missing
D93 - age started excercising: 222 missing
D102 - age when start physical labor: 86 missing 
D11C - garden work? : 8 missing
play cards, mahjong, read, watch TV



E's ADL: (1 need assistance 2 a little bit 3 none)
E1: Need assistance with bathing: 29 missing
E5 - continence: 4 misisng



F1 - Years of schooling: 25 missing
F2 - occupation: 7 missing
F31 - Main source of financial support: 



```{r}
tally(data$E5)
```

